Sparse Function-space Representation \\ of Neural Networks

### 2023-05-15 paper/main.tex
Sec. 1.1

\subsection{Related work}
\label{sec:related}
%
% General Bayesian deep learning
\textbf{Bayesian deep learning}
Probabilistic method in deep learning~\cite{Wilson:ensembles,neal1995bayesian} have recently gained increasing attention in the machine learning community as a means for uncertainty quantification (\eg, \cite{kendall2017what,wilson2020bayes}) and model selection (\eg,~\cite{immer2021scalable,antoran2022marginal}) with advancements in prior specification (\eg, \cite{cho2009kernel,meronen2020stationary,meronen2021periodic,fortuin2021bayesian,nalisnick2018do}) and efficient approximate inference under the specified model.
Calculating the posterior distribution of a Bayesian neural network is usually intractable, and approximate inference techniques need to be used, such as variational inference \cite{blei2017variational}, deep ensembles \cite{lakshminarayanan2017simple}, MC dropout \cite{gal2016dropout}, or Laplace approximation \cite{ritter2018kfac,kristiadi2020being,immer2021improving}---each having strengths and weaknesses. % point out strngths and weaknesses for each

% Introduce the laplace GGN + shortcomings
%\textbf{Function-space methods}
One common approach for uncertainty with neural networks is a Laplace-GGN approximation \citep{daxberger2021laplace}, which takes a trained neural network and linearises it around the optimal weights. The Hessian can be efficiently approximated using the generalized Gauss--Newton approximation (GGN)~\cite{botev2017practical} but typically involves a cubic scaling in the number of parameters, in practice a further approximation such as the Kronecker factorisation \cite{martens2015optimizing,ritter2018kfac} is needed. The resulting linear model can be used to obtain uncertainty estimates and refine the neural network predictions \citep{immer2021scalable}, and is linear with respect to the weights. Therefore, as shown in \cite{immer2021scalable, khan2019approximate,maddox2021fast}, the linear model can be converted to a GP, which is then used to obtain the uncertainty estimates. 
%
However, the GP introduces a cubic scaling $\mathcal{O}(N^3)$ in the number of data points $N$. Previous approaches apply crude approximations to alleviate the computational costs, \eg, by considering only a subset of the training data \cite{immer2021scalable}. In the GP community, the scaling problem is not seen as an issue any more due to efficient and stochastic methods for sparse approximations (\eg, \cite{hensman2013gaussian,wang2019exact}). However, it is not entirely clear how to combine the linearization of the neural network with sparse methods \cite{titsias2009variational} without falling back to crude subset methods. We tackle this problem by framing the problem in the dual parameters \cite{csato2002sparse} of the GP that has previously been used for non-conjugate likelihood models in GPs \cite{adam2021dual}.

% CL section
\textbf{Function-space methods for sequential learning} The central problem in continual learning is how to deal with the non-stationary nature of the training distribution, which causes the training process to overwrite the previously learnt parameters---leading to the model forgetting the previously learnt functions. 
The approaches proposed in the CL literature can be categorized into inference-based, rehearsal-based, and model-based methods; we refer the reader to \citep{parisi2019continual, de2021continual} for complete reviews. The methods in the first category usually tackle this problem with weight-space regularization, such as EWC~\citep{kirkpatrick2017overcoming}, SI~\citep{zenke17a}), or VCL~\citep{nguyen-tuongModel2009} that induce retention of previously learned information in the weights alone.
However, encoding functional knowledge in the weights does not guarantee a good quality of the predictions, thus a better to introduce a function-space regularization \citep{li2018lwf, benjamin2018measuring, titsias2019functional, buzzega2020dark, pan2020continual, rudner2022continual}. These methods bridge the gap between objective and rehearsal-based methods since they rely on a subset of the training data for each task to compute the regularization term. Recently proposed methods, such as DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual} achieve state-of-the-art performances among the objective-based techniques in several CL benchmarks.

@@

% by linearising a neural network around the Maximum a Posteriori (MAP) weights.
% a GP posterior -- can be obtained
%
In supervised learning, given a data set $\dataset = \{(\vx_{i} , \vy_{i})\}_{i=1}^{n}$, with input $\vx_i \in \inputDomain$ and output $\vy_i \in \outputDomain$ pairs, the weights $\weights \in \R^{P}$ of a neural network, $f_\mathbf{w} : \inputDomain \to \outputDomain$, are usually trained to minimize the (regularized) empirical risk,
%
\begin{equation} \label{eq-empirical-risk}
  \weights^{*} = 
  \arg \min_{\weights} \mathcal{L}(\dataset,\weights) =
  \arg \min_{\weights} \textstyle\sum_{i=1}^{n} \ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i) + \delta \mathcal{R}(\weights).
\end{equation}
%
If $\ell(f_\weights(\vx_{i}), \vy_i) = -\log(p(\vy \mid f_\weights(\vx_{i}))$ and the regularizer $\mathcal{R}(\weights) = \frac{1}{2}\|\weights\|^{2}_2$, then we can view \cref{eq-empirical-risk} as the maximum {\it a~posteriori} (MAP) solution to a Bayesian objective, where the regularization weight takes the role of a prior precision parameter, \ie, $p(\vw) = \Norm(\vzeros, \delta^{-1} \MI)$.
%Bayesian inference offers a principled approach to quantifying uncertainty in neural networks.
%The goal is to find the posterior over the weights ${p(\vw \mid \vy) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ as it represents our belief in the parameters after combining data $\dataset$ with our prior $p(\vw)$.
%Although the true posterior $p(\vw \mid \dataset)$ is intractable given the non-linearities of the neural network, one can resort to sampling techniques such as Hamiltonian Monte-Carlo (HMC).
%However, for most applications their high computational costs make them impractical so we need approximate inference techniques.
%In \cref{fig:banana}, we show a qualitative example of the induced function-space posterior obtained with our method compared to the posterior obtained through HMC sampling on the Banana toy dataset. 
The posterior over the weights ${p(\vw \mid \dataset) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ is generally not available in closed form. Sampling methods that characterize the posterior with a finite set of samples in weight-space---such as the Hamiltonian Monte Carlo (HMC) baseline used in \cref{fig:banana} (right)---are general-purpose, but comuptationally heavy and impractical for downstream applications.

>>>>

\subsection{Related work}
\label{sec:related}
%
% General Bayesian deep learning
\textbf{Bayesian deep learning}
Probabilistic methods in deep learning~\cite{Wilson:ensembles,neal1995bayesian} have recently gained increasing attention in the machine learning community as a means for uncertainty quantification (\eg, \cite{kendall2017what,wilson2020bayes}) and model selection (\eg,~\cite{immer2021scalable,antoran2022marginal}) with advancements in prior specification (\eg, \cite{cho2009kernel,meronen2020stationary,meronen2021periodic,fortuin2021bayesian,nalisnick2018do}) and efficient approximate inference under the specified model.
Calculating the posterior distribution of a Bayesian neural network is usually intractable, and approximate inference techniques need to be used, such as variational inference \cite{blei2017variational}, deep ensembles \cite{lakshminarayanan2017simple}, MC dropout \cite{gal2016dropout}, or Laplace approximation \cite{ritter2018kfac,kristiadi2020being,immer2021improving}---each with its own set of strengths and weaknesses.

%\textbf{Function-space methods}
%One common approach for uncertainty with neural networks is a Laplace-GGN approximation \citep{daxberger2021laplace}, which takes a trained neural network and linearises it around the optimal weights. The Hessian can be efficiently approximated using the generalized Gauss--Newton approximation (GGN)~\cite{botev2017practical} but typically involves a cubic scaling in the number of parameters, in practice a further approximation such as the Kronecker factorisation \cite{martens2015optimizing,ritter2018kfac} is needed. The resulting linear model can be used to obtain uncertainty estimates and refine the neural network predictions \citep{immer2021scalable}, and is linear with respect to the weights. Therefore, as shown in \cite{immer2021scalable, khan2019approximate,maddox2021fast}, the linear model can be converted to a GP, which is then used to obtain the uncertainty estimates.

% Introduce the laplace GGN + shortcomings
\textbf{Function-space methods}
Function-space perspectives on uncertainty in neural networks often use a Laplace-GGN approximation \citep{daxberger2021laplace}, linearizing a trained neural network around optimal weights. The Hessian can be efficiently approximated using the generalized Gauss--Newton approximation (GGN)~\cite{botev2017practical}. While efficient, it suffers from cubic scaling in parameter count, necessitating approximations like Kronecker factorisation \cite{martens2015optimizing, ritter2018kfac}. This linear model, linear in the weights, refines predictions and provides uncertainty estimates \citep{immer2021scalable}. This model has a convenient interpretation as a GP \cite{immer2021scalable, khan2019approximate, maddox2021fast}.
%
However, the GP's cubic $\mathcal{O}(N^3)$ scaling with data points $N$ requires costly approximations, often resorting to using data subsets \cite{immer2021scalable}. In the GP community, sparse approximations have mitigated this scaling issue (\eg, \cite{titsias2009variational,hensman2013gaussian}), but combining neural network linearization with sparse methods remains unclear without resorting to subset methods. Our work addresses this by utilizing the GP's dual parameters \cite{csato2002sparse}, previously applied to non-conjugate likelihood models \cite{adam2021dual}.

%wang2019exact
%
%However, the GP introduces a cubic $\mathcal{O}(N^3)$ scaling in the number of data points $N$. Previous approaches apply crude approximations to alleviate the computational costs, \eg, by considering only a subset of the training data \cite{immer2021scalable}. In the GP community, the scaling problem is not seen as an issue any more due to efficient and stochastic methods for sparse approximations (\eg, \cite{hensman2013gaussian,wang2019exact}). However, it is not entirely clear how to combine the linearization of the neural network with sparse methods \cite{titsias2009variational} without falling back to crude subset methods. We tackle this problem by framing the problem in the dual parameters \cite{csato2002sparse} of the GP that has previously been used for non-conjugate likelihood models in GPs \cite{adam2021dual}.

% CL section
%\textbf{Function-space methods for sequential learning} The central problem in continual learning is how to deal with the non-stationary nature of the training distribution, which causes the training process to overwrite the previously learnt parameters---leading to the model forgetting the previously learnt functions. 
%The approaches proposed in the CL literature can be categorized into inference-based, rehearsal-based, and model-based methods; we refer the reader to \citep{parisi2019continual, de2021continual} for complete reviews. The methods in the first category usually tackle this problem with weight-space regularization, such as EWC~\citep{kirkpatrick2017overcoming}, SI~\citep{zenke17a}), or VCL~\citep{nguyen-tuongModel2009} that induce retention of previously learned information in the weights alone.
%However, encoding functional knowledge in the weights does not guarantee a good quality of the predictions, thus a better to introduce a function-space regularization \citep{li2018lwf, benjamin2018measuring, titsias2019functional, buzzega2020dark, pan2020continual, rudner2022continual}. These methods bridge the gap between objective and rehearsal-based methods since they rely on a subset of the training data for each task to compute the regularization term. Recently proposed methods, such as DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual} achieve state-of-the-art performances among the objective-based techniques in several CL benchmarks.

\textbf{Function-space methods for sequential learning} 
Continual learning (CL) hinges on how to deal with the non-stationary training distribution, wherein the training process may overwrite previously learned parameters, causing catastrophic forgetting. CL approaches fall into inference-based, rehearsal-based, and model-based methods (see \cite{parisi2019continual, de2021continual} for an overview). The methods in the first category usually tackle this problem with weight-space regularization, such as EWC~\citep{kirkpatrick2017overcoming}, SI~\citep{zenke17a}), or VCL~\citep{nguyen-tuongModel2009} that induce retention of previously learned information in the weights alone. However, encoding functional knowledge in the weights does not guarantee good quality of the predictions. Function-space regularization techniques~\cite{li2018lwf, benjamin2018measuring, titsias2019functional, buzzega2020dark, pan2020continual, rudner2022continual} address this by relying on training data subsets for each task to compute the regularization term. These methods bridge the gap between objective and rehearsal-based methods since they use a subset of the training data for each task to compute the regularization term. Recently proposed methods, \eg DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual} achieve state-of-the-art performance among the objective-based techniques in several CL benchmarks.

%%

@@

% by linearising a neural network around the Maximum a Posteriori (MAP) weights.
% a GP posterior -- can be obtained
%
In supervised learning, given a data set $\dataset = \{(\vx_{i} , \vy_{i})\}_{i=1}^{N}$, with input $\vx_i \in \inputDomain$ and output $\vy_i \in \outputDomain$ pairs, the weights $\weights \in \R^{P}$ of a neural network, $f_\mathbf{w} : \inputDomain \to \outputDomain$, are usually trained to minimize the (regularized) empirical risk,
%
\begin{equation} \label{eq-empirical-risk}
  \weights^{*} = 
  \arg \min_{\weights} \mathcal{L}(\dataset,\weights) =
  \arg \min_{\weights} \textstyle\sum_{i=1}^{N} \ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i) + \delta \mathcal{R}(\weights).
\end{equation}
%
If $\ell(f_\weights(\vx_{i}), \vy_i) = -\log(p(\vy \mid f_\weights(\vx_{i}))$ and the regularizer $\mathcal{R}(\weights) = \frac{1}{2}\|\weights\|^{2}_2$, then we can view \cref{eq-empirical-risk} as the maximum {\it a~posteriori} (MAP) solution to a Bayesian objective, where the regularization weight takes the role of a prior precision parameter, \ie, $p(\vw) = \Norm(\vzeros, \delta^{-1} \MI)$.
%Bayesian inference offers a principled approach to quantifying uncertainty in neural networks.
%The goal is to find the posterior over the weights ${p(\vw \mid \vy) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ as it represents our belief in the parameters after combining data $\dataset$ with our prior $p(\vw)$.
%Although the true posterior $p(\vw \mid \dataset)$ is intractable given the non-linearities of the neural network, one can resort to sampling techniques such as Hamiltonian Monte-Carlo (HMC).
%However, for most applications their high computational costs make them impractical so we need approximate inference techniques.
%In \cref{fig:banana}, we show a qualitative example of the induced function-space posterior obtained with our method compared to the posterior obtained through HMC sampling on the Banana toy dataset. 
The posterior over the weights ${p(\vw \mid \dataset) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ is generally not available in closed form. Sampling methods that characterize the posterior with a finite set of samples in weight-space---such as the Hamiltonian Monte Carlo (HMC) baseline used in \cref{fig:banana} (right)---are general-purpose, but comuptationally heavy and impractical for downstream applications.
